### 2018-12-27 revisions/review-responses.tex

        %---------------------------------------------------------------------
        \question  Table 5: Please provide some comments on the uncertainties - 
        where do they come from? Also, the ``reference'' results need to state 
        whether ``initial'' or ``equilibrium''
        \begin{solution}
                In the Table 5 added information that reference column contains data for initial 					fuel salt composition. Details about uncertainties has been added:

                Uncertainty for each temperature coefficient was obtain by propagating 			  				statistical error for effective multiplication factor calculation (confidence
                 interval $\pm$1.9 pcm) provided by SERPENT2 and also appears in Table 5.
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 26, line 425: ``Relatively large'' compared with what? 
        Perhaps results for an LWR case would be a good basis for comparison?
        \begin{solution}
                Sentence has been extended as follows:

                ...the total temperature coefficient of reactivity is relatively large and
                 negative during reactor operation (comparing with conventional PWR which 							temperature coefficient about -1.71 pcm/$^\circ$F $\approx$ -3.08 pcm/K 							\cite{forget_integral_2018}), despite positive MTC, and affords excellent 
                 reactor stability and control.
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 27, section 3.8: It needs to be made more clear that 
        these results were calculated, and that they are taken from the code 
        output.
        \begin{solution}
                This has now been clarified in the text:

                Table 7 summarizes the six factors for both initial and 
				equilibrium fuel salt composition. These factors and their statistical 								uncertainties have been calculated using SERPENT2 code for initial fuel salt 
				composition (see Table 2) and for equilibrium salt composition which was 
				obtained with SaltProc. 
        \end{solution}

### 2018-12-28 revisions/review-responses.tex

		        The neutron energy spectrum at equilibrium is harder than at startup due
		         to plutonium and other strong absorbers accumulation in the core during 
		         reactor operation.  
        \end{solution}

        %---------------------------------------------------------------------
        \question  Fig 12: units on y-axis?
        \begin{solution}
                Thanks for catching this. Units $\frac{n}{cm^2 s}$ has been added.
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 24, line 389: Should that be ``233U'' and not ``233Th''?
        \begin{solution}
                Yes, we meant $^{233}$U production. The typo has been fixed.
        \end{solution}

        %---------------------------------------------------------------------
        \question  Table 5: Please provide some comments on the uncertainties - 
        where do they come from? Also, the ``reference'' results need to state 
        whether ``initial'' or ``equilibrium''
        \begin{solution}
                In the Table 5 added information that reference column contains data for initial 					fuel salt composition. Details about uncertainties have been added:

                 Uncertainty for each temperature coefficient was obtain by 
                 propagating statistical error of effective multiplication factor 
                 calculation provided by SERPENT2 and also appears in Table 5. 
                 Other sources of uncertainty, such as cross section libraries 
				 uncertainty, error in salt and graphite density correlations, 
				 are not treated here.
        \end{solution}

@@

        %---------------------------------------------------------------------

        \question Can the proposed tool adjust reactivity?
        \begin{solution}
                <++>
        \end{solution}

        %---------------------------------------------------------------------

        \question ``The main physical principle underlying the reactor 
        temperature feedback is an expansion of material that is heated''. This 
        sentence would deserve some support data. Can you please calculate 
        separately the effect of temperature (Doppler in fuel and spectral 
        shift in graphite) and density?
        \begin{solution}
                Thank you for the excellent recommendation. Effects of temperature
                 and density have been calculated separately and added in Table 5.
                 Moreover, during these simulations we have discovered mistake in 
                 fuel salt and graphite density correlations and completely 
                 recalculated temperature coefficients of reactivity with lower 
                 statistical error. Now initial total temperature coefficient is 
                 closer to the reference and statistical uncertainty was reduced 
                 from 0.046 to 0.038 pcm/K.
        \end{solution}

@@

        %---------------------------------------------------------------------

        \question How uncertainties have been calculated in Table 5? Is this 
        just statistical uncertainty from Serpent calculations? Please clarify
        \begin{solution}
                Yes, uncertainties were determined from statistical error 
                from SERPENT output. Following passage has been added to clarify
                 this point:

                 Uncertainty for each temperature coefficient was obtain by 
                 propagating statistical error of effective multiplication factor 
                 calculation provided by SERPENT2 and also appears in Table 5. 
                 Other sources of uncertainty, such as cross section libraries 
				 uncertainty, error in salt and graphite density correlations, 
				 are not treated here.
        \end{solution}

@@

        %---------------------------------------------------------------------

        \question For calculating the coefficients in table 5, are you only 
        changing densities, or also dimensions? Please clarify and provide a 
        justification.  
        \begin{solution}
        		We have changed both densities and dimensions. This passage 
        		has been modified to read:

                A new geometry input for SERPENT2, which takes into account 
                displacement of graphite surfaces, was created based on this information.
                 For the displacements calculation it was assumed that the interface 
                 between graphite reflector and vessel did not move, and the vessel 
                 temperature did not change. This is the most reasonable assumption for
				 the short-term reactivity effects because inlet salt is cooling graphite 
				 reflector and inner surface of the vessel.
        \end{solution}

@@

        %---------------------------------------------------------------------

        \question ``The fuel temperature coefficient (FTC) is negative for both 
        initial and equilibrium fuel compositions due to thermal Doppler 
        broadening of the resonance capture cross sections in the thorium.'' 
        What is the effect of density?
        \begin{solution}
                This passage has been modified to illuminate the effect of density:

				The fuel temperature coefficient (FTC) is negative for both initial and 
				equilibrium fuel compositions due to thermal Doppler broadening of the 
				resonance capture cross sections in the thorium. Small positive effect 
				of fuel density on reactivity is increasing from +1.21 pcm/K for the 
				reactor startup to +1.66 pcm/K for equilibrium fuel composition which 
				has negative effect on FTC magnitude during the reactor operation. 
        \end{solution}

### 2018-12-28 results.tex
Reviewer #1 comments all implmented #32
Table 5 thermal coefficients recalculated.Resolve issue #31

Uncertainty for each temperature coefficient also appears in 
Table~\ref{tab:tcoef}. The main physical principle underlying the reactor 

@@

   Reactivity coefficient [pcm/K]  & Initial      & Equilibrium  & Reference 
        \cite{robertson_conceptual_1971} \\  \hline
Fuel salt        & $-3.22\pm0.044$ & $-1.53\pm0.046$ & $-3.22$  \\
Moderator        & $+1.61\pm0.044$ & $+0.97\pm0.046$ & $+2.35$  \\
Total            & $-3.1\pm0.04$   & $-0.97\pm0.046$ & $-0.87$  \\ \hline

@@

geometry input was created based on this information.

The fuel temperature coefficient (FTC) is negative for both initial and 
equilibrium fuel compositions due to thermal Doppler broadening of the resonance 
capture cross sections in the thorium. This is in good agreement with earlier 
research \cite{robertson_conceptual_1971,park_whole_2015}. The moderator 
temperature coefficient (MTC) is positive for the startup composition and decreases 
during reactor operation because of spectrum hardening with fuel depletion. 
Finally, the total temperature coefficient of reactivity is negative for both 
cases, but decreases during reactor operation due to spectral shift. In 
summary, even after 20 years of operation the total temperature coefficient of 
reactivity is relatively large and negative during reactor operation, despite 
positive MTC, and affords excellent reactor stability and control.

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. The non-leakage probability for both fast 
and thermal neutrons does not change during reactor operation because these 
values are not largely affected by the neutron spectrum shift. In contrast, 
neutron reproduction factor ($\eta$), resonance escape probability (p), and 
fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. For example, the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to batch-wise 
removal of strong absorbers (Rb, Sr, Cs, Ba). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}

>>>>

Uncertainty for each temperature coefficient was obtain by propagating statistical 
 error of effective multiplication factor calculation 
provided by SERPENT2 and also appears in 
Table~\ref{tab:tcoef}. Other sources of uncertainty, such as cross section libraries 
uncertainty, error in salt and graphite density correlations, are not treated here.
 The main physical principle underlying the reactor 

@@

   Reactivity coefficient [pcm/K]  & Initial      & Equilibrium  & Reference \qquad (initial)\cite{robertson_conceptual_1971} \\  \hline
Doppler in fuel salt   			& $-4.73\pm0.038$ & $-4.69\pm0.038$ & $-4.37$  \\
Fuel salt density      			& $+1.21\pm0.038$ & $+1.66\pm0.038$ & $+1.09$  \\
Total fuel salt        			& $-3.42\pm0.038$ & $-2.91\pm0.038$ & $-3.22$  \\ \hline
Graphite spectral shift    		& $+1.56\pm0.038$ & $+1.27\pm0.038$ &          \\
Graphite density   				& $+0.14\pm0.038$ & $+0.23\pm0.038$ &          \\
Total moderator (graphite)		& $+1.69\pm0.038$ & $+1.35\pm0.038$ & $+2.35$  \\ \hline
Total core      				& $-1.64\pm0.038$ & $-1.58\pm0.038$ & $-0.87$  \\ \hline

@@

geometry input for SERPENT2, which takes into account displacement of graphite 
surfaces, was created based on this information. For the displacements calculation 
it was assumed that the interface between graphite reflector and vessel did not move,
 and the vessel temperature did not change. This is the most reasonable assumption for
 the short-term reactivity effects because inlet salt is cooling graphite reflector and 
inner surface of the vessel.

The fuel temperature coefficient (FTC) is negative for both initial and 
equilibrium fuel compositions due to thermal Doppler broadening of the resonance 
capture cross sections in the thorium. Small positive effect of fuel density on 
reactivity is increasing from +1.21 pcm/K for the reactor startup to +1.66 pcm/K for 
equilibrium fuel composition which has negative effect on FTC magnitude during the 
reactor operation. This is in good agreement with earlier 
research \cite{robertson_conceptual_1971,park_whole_2015}. The moderator 
temperature coefficient (MTC) is positive for the startup composition and decreases 
during reactor operation because of spectrum hardening with fuel depletion. 
Finally, the total temperature coefficient of reactivity is negative for both 
cases, but decreases during reactor operation due to spectral shift. In 
summary, even after 20 years of operation the total temperature coefficient of 
reactivity is relatively large and negative during reactor operation (comparing 
with conventional PWR which has temperature coefficient about -1.71 pcm/$^\circ$F 
$\approx$ -3.08 pcm/K \cite{forget_integral_2018}), despite positive MTC, and 
affords excellent reactor stability and control.

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. These factors and their statistical uncertainties
 have been calculated using SERPENT2 code for initial fuel salt composition (see 
Table~\ref{tab:msbr_tab}) and for equilibrium salt composition which was obtained 
with SaltProc. The non-leakage probability for both fast 
and thermal neutrons does not change during reactor operation because these 
values are not largely affected by the neutron spectrum shift. In contrast, 
neutron reproduction factor ($\eta$), resonance escape probability (p), and 
fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. For example, the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to batch-wise 
removal of strong absorbers (Rb, Sr, Cs, Ba). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}
